Speculation in foreign exchange: noise or information

### 2015-07-13 Knitr/NewPaper.Rnw
Read through complete.  One more read and then send. Appendix still does not work correctly.

Towards the other end of the spectrum, Milton Friedman argued that speculation was a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium. In arguing in favour of flexible exchange rates,  \citet{FriedmanPositive} asserted that successful speculators would be those who managed to correctly buy when prices  were below fundamental value and sell above that price. Therefore, Friedmand argued, the speculative process would favour those able to access or utilise information over the uninformed and, through a Darwinian process, informed speculation would dominate.  This view regards speculation as an key lubricant of financial markets as it provides liquidity and facilitates the process of price discovery.

The aim of this paper is to understand more about the nature of speculation by investigating the relationship between both the intensity of speculation and the weight of speculative positions with the movement of prices in the foreign exchange market.   If speculators are uninformed and their activity can be regarded as noise, the most extreme sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place; if speculators are informed prices should follow a random walk after either of these extremes.  
    The foreign exchange market is utilised because it is the largest and most liquid financial market in the world.  The latest survey taken by the Bank for International Settlements (BIS) in April 2013 estimated average daily turnover at \$5.4trn with average daily spot transactions of \$2.0trn See \citet{BISFX2013}.  The cost of trading in this market, when measurd in terms of the size of the bid-ask spread, is extremely low. An analysis by \citet{Steely2013} of the bid-ask spreads for USD-JPY, GBP-USD and EUR-USD for the period January 2001 to 2005 and finds that the average spread for each is 0.342\%, 0.305\% and 0.326\% respectively.  This would be the cost for a round-trip (buying and selling immediately).

A unique set of option prices are used to identify the intensity of speculative sentiment.  US regulatory positions measure the prominence of speculators in the market. Speculation does not appear to be random, there is momentum to the rise and fall in sentiment and positions.  However, while currency prices co-move with sentiment and positions, there is no informational content from either: knowing the the strength of speculative opinion or the weight of speulative positions does not improve forecasts of future returns. This supports the suggestion that speculative activity has a permanent affect on prices and that is consistent with speculation being part of the process of price discovery. Speculators do not seem to drive prices from fundamental value and their presence does not seem to hamper the incorporation of information into prices.

@@

One key element of the microstructure literature involves distinguishing between uninformed or noise-traders and those who have infomration. The \citet{Kyle1985Continuous} model has market-makers reciving orders from the informed and the uninformed.  On the assumption that the uniformed arrive in a random fashion without any reason to trade, the weight of orders implies something about the activity of the informed and causes prices to change in the process of price discovery.  The bid-ask spread is the payment that the market-maker recieves in return for taking the risk of meeting an informed trader.  Ultimately, the uninformed traders make payments to the informed traders with the mark-makers taking a cut for the risk and the privision of exchange.  See also \citet{glosten1985bid}.

@@

\citet[p. 1]{GortonNoise} categorise noise-traders as all those trading in markets for ``non-information-based reasons''. In this case, uninforrmed speculation provides a rational and an opportunity for the trading of financial securities. If all economic agents know the fundamental value of a security how can some be buying and some be selling? There must be different opinions.  Why will traders be prepared to pay the cost of acquiring information if they are unable to take advantage of that information? With speculative noise-traders, the informed can take advantage of their information.  This creates a justification for acquiring information. Speculative noise-traders can therefore create a gap between the price and the fundamtental value that will encourage informed traders to pay the cost of acquiring information. The divergence between this price and fundametnal value is thereby essential to encourage te process of price discovery. 
%
 \citet[p.529]{BlackNoise} suggested that this combination was essential to the smooth function of financial markets. It is ``the arbitrary element in expectations''. Therefore, noise makes financial markets possible but imperfect: the more noise-traders that there are, the more liquidity and the easier it is to trade; the more noise-traders, the higher the level of inefficiency in the market and the more likely that price will be driven from fundamental value.

@@

However, though the idea that noise is a random variable may be a construction that is useful for modelling and it may be justified in analysis of broader themes of microstructure and levels of market efficiency, this exogenous construct may be an over-simplification that is inapprpropriate when looking for more understanding of speculation itself.  Institutional and psychological explanations for noise-trading frequently suggest that there are endogenous non-random elements.

The institutional explanations are assciated with the allocation of credit for investment and trading purposes:  as asset values decline exchanges, banks and other providers of liqudiity require larger margins and deeper haircuts; investors liquidate their fund holdings; the atttraction of hedging increases.  See \citet{BrunnermeierLiquidity} and \citet{Gorton} for an overview of the institutional literature.  Psychological explanations revolve around the common panic, the group-thining and the tendency to read private information from the actions of other investors: the so-called \emph{wisdom of crowds} depends on independent evaluation but as Keynes noted in his discussion of the \emph{beauty contest}, it is generally more important to identify what other investors will believe than acertain underlying, fundamental value \citet[p.101]{Keynes1936}

it has also been suggested that speculative technieques can change overtime.  For example, \citet{ShillerFashion} discusses the use of fashionable techniques or fads in investment.  In this way, speculation noise is associated with sentiment and measures of sentiment are then used as ways of measuring speculation. 
%It would be good to have some more modern ideas here.  Maybe they are just the empirical results. 
%this paragraph is not very good.  It needs to be updated.  Look at the file in citulike. 
\subsection{Empirical work}
Empirical assessment of speuclation has been limited by the scarcity of data.  It is not very easy to distinguish between speculators and other types of investors.  It is even harder to distinguish those that may be noise-traders from those assumed to be informed.  One approach has been to assume that noise-traders utilise simple investment strategies and to use tests of these strategies as a means of testing speculation.  For example speculation can be identified with momentum trading. It is also possible to assume that certain investors, such as investment banksa and hedge funds, are informed while other trade without inforation. The latter group may include day-traders or private individuals.

@@

The debate about a tax on financial transactions, initiated by \citet{TobinTax} and now called the \emph{the Tobin tax}, has been intended to reduce speculation.  The surge in commodities prices that was seen prior to the GFC was seen to be partially the consequence of increased speculation in commodities and the creation of exchange traded funds (ETF) focusing on commodities markets.  The argument for a tax on finanical transactions depends on the idea that speculators will drive the price away from its fundamental value.

@@

\section{Methodology}
The aim of this study is to test whether the noise-trader model provides an appropriate explanation for speculator activity in the  foreign exchange market. This will be use to gain a fuller understanding of speculative behaviour.   The experiement will involve three steps:  assess the price implication of the noise-trader model;  measure speculative activity in the foreign exchange market; finally, assess whether extreme speculative activity in the foreign exchange market is consistent with the noise-trader model.  If the speculative activity that is seen in the foreign exchange market produces results that are like the noise-trader model, it can be said that speculation is noise trading as Keynes asserted, if not it may be more informed as suggested by Friedman.  
% where it is widely perceived that there are a relatively large number of traders engaged in technical analysis.  See \citet gc{FrootChartists} for a survey.  %this may go in the explanation for why fx is used.  Use this elsewhere. Add more?

@@

The following four effects are noted by \citet[pp. 14-15]{Delong1990noise}:  the \emph{Friedman Effect}, as informed speculators move prices back towards fundamental value; \emph{Price Pressure} as the speculative noise-traders drive prices towards their misperceived vaue; \emph{Hold More} as noise-traders misperceive the value of the risky asset and tend to hold more of it; \emph{Create Space} is the noise-trader risk caused by the difference between their perception and the actual value, it is an added risk that is caused by the variability of noise-trader belief as noise-traders could drive the price even further from its fundamental value. In this model, tHe more numerous the noise-traders, the more volatile prices.

The Friedman-effect and the Price-pressure are the effects that are studied here.  Their combination suggests that when the weight of noise-traders is high or the times when they have particularly stong opinions will tend to be those when there is the greatest risk of reversals.  The hold-more effect suggests that the weight of speculative involvement in the market will tend to increase as a trend.  Create-space effect would suggest that when the weight of noise-traders is high, there is incresaed risk.  %this can be assessed.

@@

To understand what the noise-trader model of speculative behaviour means for foreign exchange rates, a \emph{Monte Carlo} simulation is carried out with a variety of mis-perceptions and noise-trader weights.  Two time series are created.  The first series is based on a random noise-trade misperception, $(\rho^*)$.  This is called the Misperception Model (MPM). The second is based on a random weight of noise-traders $(\mu)$ and is called the Random Weight Model (RWM).   The simulation creates a 1000 period price series from Equation \ref{eq:NP2}. The dividend paid on the risky asset $(r)$ is posted at 10\%; the risk-aversion parameter $(\gamma)$ is assumed to be 0.3 (in line with that suggested in \citet[p. 127]{FrobesBehaviour}); starting from a price of 100, the average noise-trader mis-perception $(p^*)$ is a random variable in the first simulation with a mean of 4 and a variance $(\sigma_{\rho^*}^2)$ of 2, and has a fixed value of 4 in the second simulation; the weight of noise-traders $(\mu)$ is 50\% in the first case but is modelled as a two-period moving average of a normal random variable centred at 50\% with a standard deviation of 20\% and equally weighted average of its previous value (starting at 50\%) to allow smooth adjustment, in the second simulation.   
% it is possible to have other specifications. For example, the noise-traders oculd become more prominent if their strategy is more successful,

A 1000 period prices simulation is created for each model and the returns for these series are calculated as the log change in price.  Those periods when mis-perception or the weight of speculators is \emph{extreme} are identified as those that are at or above the ninety-fifth percentile.  These extremes are the \emph{event} for an event study. The R package \emph{eventstudies} createed by \citet{eventstudies} is used for the data manipulation.  Descriptive statistics for the whole series of returns and for those when either the extent of their misperception or the weight of noise-traders is extreme are presented in Table \ref{tabref:MC1}.  The process is then repeated for one thousand simulations of each model so that confidence intervals can be established for the estimates of the descriptive statistics.

@@

Table \ref{tabref:DS1} in Appendix \ref{apref:1} shows descriptive statistics for the exchange rate data used in the study.  There are daily data for the period between January 1996 and September 2002.  This is the period for which risk-reversal data are available.  The series are of sightly different lengths as there are some gaps in the risk reversal and exchange rate data as a result of holidays and market closures.  From the raw data continuously compounded returns are calculated and multiplied by 100.  The first thing of note is that the mean return does not appear to be statistically different from zero.  This is what would be expected if, as has been identified in many studies, exchange rates follow a random walk. 
\begin{landscape}
\begin{table}[t!]
\begin{threeparttable}
\caption{Descriptive Statistics for Risk Reversals: 2 January 1996 to 6 September 2002}
\begin{tabular}{l | r r  r r r r r r r r}
  \hline
 Statistic &  AUD & EURGBP & EURSEK & GBP & CHF & EURCHF & EURJPY & EUR & CAD & JPY \\
 \hline
Number & 1702 & 1697 & 1719 & 1691 & 1374 & 1339 & 1697 & 1692 & 1701 & 1697 \\ 
  Mean & -0.37 & 0.15 & 0.71 & -0.08 & -0.28 & -0.38 & -0.48 & 0.19 & 0.12 & -0.74 \\ 
  Median & -0.30 & 0.10 & 0.80 & -0.10 & -0.15 & -0.30 & -0.10 & 0.10 & 0.15 & -0.55 \\ 
  Standard Deviation & 0.38 & 0.37 & 0.51 & 0.33 & 0.49 & 0.36 & 1.02 & 0.49 & 0.25 & 1.02 \\ 
 Max & 0.45 & 1.72 & 1.90 & 1.00 & 1.05 & 0.50 & 1.70 & 2.05 & 1.45 & 2.05 \\ 
 Min & -1.70 & -0.80 & -0.30 & -1.35 & -2.05 & -1.80 & -4.03 & -1.15 & -0.65 & -4.00 \\ 
 Skewness & -0.50* & 1.27* & -0.23* & 0.33* & -0.79* & -0.58* & -0.93* & 0.68* & 0.23* & -0.51* \\ 
  Kurtosis & 0.02 & 2.29* & -1.67 & 0.72 & 0.69 & 0.19 & 0.15 & 0.96 & 1.83 & -0.03 \\ 
      \hline
 \label{tabref:DS2}
\end{tabular}
\begin{tablenotes}
\small
\item Daily risk reversal skew.  Risk reversal is the premium of implied volatility for calls relative to the equivalent put (equal distance to the strike price, 25 delta in this case).  A figure above zero means that calls are more expensive (in volatility and price terms) than puts.  A figure below zero means that puts are more expensive. The asterisk (*) means that the estimate is significant at the 5\% level using t-statistics for the mean and the standard error of skew and standard error of excess-kurtosis respectively.   These are calculated as $SES = \sqrt{\frac{6n(n-1)}{(n-2)(N+1)(N+3)}}$ and $SEK = \sqrt{\frac{n^2-1}{(n-3)(N+5)}}$, with the assumption that this estimate is distributed normally, the skew or kurtosis is considered significant if it is more than two standard errors from zero.
\end{tablenotes}
\end{threeparttable}
\end{table}
\end{landscape}

@@

\begin{landscape}
\begin{table}[t]
\begin{threeparttable}
\caption{Descriptive Statistics for S1 (Speculative Sentiment) and S2 (Speculative Weight): 30 September 1998 to 31 December 2008}
\begin{tabular}{l | r r r r r r r  r r r }
  \hline
 Statistic &  \multicolumn{2}{c}{EUR} & \multicolumn{2}{c}{JPY} & \multicolumn{2}{c}{GBP} & \multicolumn{2}{c}{CHF} & \multicolumn{2}{c}{CAD}  \\
 & S1 & S2 & S1 & S2 & S1 & S2 & S1 & S2 & S1 & S2 \\  
\hline
Number & 522 &  522 & 849 & 849 & 849 & 849 & 847 & 847 & 849 & 849\\
Mean &  31.31 & 43.43 & -23.90 & 44.21 & 2.76 & 41.68 & -21.14 & 46.13 & 6.68 & 41.23\\
Median & 42.75 & 39.56 & -33.30 & 43.56 & 11.15 & 38.88 & -37.32 & 42.88 & 7.44 & 39.81\\
Standard deviation  & 41.94 & 16.13 & 51.49 & 12.17 & 60.59 & 16.46 & 60.76 & 15.73 & 55.26 & 15.33\\
Max & 95.67 & 86.06 & 87.43 & 84.50 & 96.15 & 89.45 & 100.00 & 92.49 & 100.00 & 86.63\\
Min &  -100.00 & 11.15 & -99.83 & 12.88 & -100.00 & 9.10 & -100.00 & 14.61 & -90.08 & 11.89\\
Skew  & -0.88* & 0.55* & 0.33* & 0.57* & -0.17 & 0.76* & 0.45* & 0.65* & 0.02 & 0.50*\\
Kurtosis & 0.18 & -0.57 & -1.14 & 0.64 & -1.45 & 0.15 & -1.25 & -0.15 & -1.35 & -0.28\\
\hline
 \label{tabref:DS4}
\end{tabular}
\begin{tablenotes}
\small
\item S1 is the first sentiment indicator, calculated as ratio of net non-commercial long less non-commercial short to the sum of non-commercial long and short positions multiplied by one hundred.  S2 is the second sentiment indicators, calculated as the ratio of net non-commercial long to net non-commercial short plus spreading to open interest.   An asterisk (*) means that the estimate is significantly different from zero at the 5\% level using t-statistics for the mean and the standard error of skew and standard error of excess-kurtosis respectively.  These are calculated as $SES = \sqrt{\frac{6n(n-1)}{(n-2)(N+1)(N+3)}}$ and $SEK = \sqrt{\frac{n^2-1}{(n-3)(N+5)}}$, with the assumption that this estimate is distributed normally, the skew or kurtosis is considered significant if it is more than two standard errors from zero.
\end{tablenotes}
\end{threeparttable}
\end{table}
\end{landscape}

\subsection{Analysis of events and abnormal returns}
The aim of the study is to assess the effect of extreme speculative positions on future returns to ascertain the informational content of these extremes.  Therefore, the events that will be used to assess abnormal returns are the extreme readings on sentiment and the extreme presence of speculators.   Any reading above the ninety-ninth percentile or below the first percentile is deemed to be extreme.  For the risk reversal skew, as a secondary test, a reading above the ninety-fifth and below the fifth percentile is also established as an event.   The Event Studies R package was used to manage the data see \citet{eventstudies} for fuller details on the package and its use.  The cumulative abnormal returns for the whole of the event window (CARW) and the cumulative abnormal returns for after the event (CARA) were recorded.  These cumulations are just the summation of the continuously compounded returns over the appropriate period.

@@

What is more interesting is whether knowledge about the extreme reading at time $t = 0$ provides any information about future returns.  If it is known that speculative sentiment is at the extreme, does this mean that it is more likely that there will be a reversal? Is there an inefficiency?  This information is provided in columns 5, 8, 11 and 14 of Table \ref{tabref:RR1}. A reversal is indicated by a significant negative reading for highs or positive reading for lows.   Looking, for example, at the four-day window in column 8, it is clear that the extreme risk reversal skew does not provide any information about future returns.  Of the eighteen cases using the 95\% th percentile to identify an extreme, ten show a continuation, rather than a reversal, that is not significantly different from zero; one shows a significant continuation; five show reversals that are not significant; only two show significant reversal (EURCHF in each case).  For the more extreme event when the 99\% the percentile is used to measure the extreme, there is an even more mixed picture.  Just over half the cases with the four-day window show a reversal (8 of 18), but only 2 of these are statistically significant.

\begin{landscape}

@@

\begin{table}[t]

@@

\end{table}
\end{landscape}

@@

Once again, there is a very strong association between returns and the movement of speculators.  For the 4-day window and the S1 and S2 measures, every one of the currencies studied, high and low, had a mean cumulative return for the whole of the event window that was in the same direction as sentiment.  When re-sampled a thousand times, more than ninety five percent of the means calculated were greater than zero in all cases but one (GBP).  Once again there is evidence here of momentum and of speculator positions following the movement of prices.

However, there is little evidence that these speculators are driving prices to extremes with their extreme behaviour.  Sixty percent of the S1 cases still show a significant abnormal return in the direction of the extreme 4 weeks after the extreme has been reached; only thirty percent of the cases show a reversal and in only one of those do ninety five percent of the re-sampled means remain above zero.  For S2 it is even more extreme, with ninety percent of the cases showing a continuation and forty percent of them being significant.  Again the evidence is consistent with the gradual adjustment of prices to information.

>>>>

Towards the other end of the spectrum, Milton Friedman argued that speculation was a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium. In arguing in favour of flexible exchange rates,  \citet{FriedmanPositive} asserted that successful speculators would be those who managed to correctly buy when prices  were below fundamental value and sell above that price. Therefore, Friedman argued, the speculative process would favour those able to access or utilise information over the uninformed and, through a Darwinian process, informed speculation would dominate.  This view regards speculation as an key lubricant of financial markets as it provides liquidity and facilitates the process of price discovery.

The aim of this paper is to understand more about the nature of speculation by investigating the relationship between both the intensity of speculation and the weight of speculative positions with the movement of prices in the foreign exchange market.   If speculators are uninformed and their activity can be regarded as noise, the most extreme sentiment and the time when speculators have the greatest weight in the market should correspond with the greatest deviation of prices from fundamental value, when reversals are most likely to take place; if speculators are informed prices should follow a random walk after either of these extremes.

The foreign exchange market is utilised because it is the largest and most liquid financial market in the world.  The latest survey taken by the Bank for International Settlements (BIS) in April 2013 estimated average daily turnover at \$5.4trn with average daily spot transactions of \$2.0trn See \citet{BISFX2013}.  The cost of trading in this market, when measured in terms of the size of the bid-ask spread, is extremely low. An analysis by \citet{Steely2013} of the bid-ask spreads for USD-JPY, GBP-USD and EUR-USD for the period January 2001 to 2005 and finds that the average spread for each is 0.342\%, 0.305\% and 0.326\% respectively.  This would be the cost for a round-trip (buying and selling immediately).

A unique set of option prices are used to identify the intensity of speculative sentiment.  US regulatory positions measure the prominence of speculators in the market. Speculation does not appear to be random, there is momentum to the rise and fall in sentiment and positions.  However, while currency prices co-move with sentiment and positions, there is no informational content from either: knowing the the strength of speculative opinion or the weight of speculative positions does not improve forecasts of future returns. This supports the suggestion that speculative activity has a permanent affect on prices and that is consistent with speculation being part of the process of price discovery. Speculators do not seem to drive prices from fundamental value and their presence does not seem to hamper the incorporation of information into prices.

@@

One key element of the microstructure literature involves distinguishing between uninformed or noise-traders and those who have information. The \citet{Kyle1985Continuous} model has market-makers receiving orders from the informed and the uninformed.  On the assumption that the uniformed arrive in a random fashion without any reason to trade, the weight of orders implies something about the activity of the informed and causes prices to change in the process of price discovery.  The bid-ask spread is the payment that the market-maker receives in return for taking the risk of meeting an informed trader.  Ultimately, the uninformed traders make payments to the informed traders with the mark-makers taking a cut for the risk and the provision of exchange.  See also \citet{glosten1985bid}.

@@

\citet[p. 1]{GortonNoise} categorise noise-traders as all those trading in markets for ``non-information-based reasons''. In this case, uninformed speculation provides a rational and an opportunity for the trading of financial securities. If all economic agents know the fundamental value of a security how can some be buying and some be selling? There must be different opinions.  Why will traders be prepared to pay the cost of acquiring information if they are unable to take advantage of that information? With speculative noise-traders, the informed can take advantage of their information.  This creates a justification for acquiring information. Speculative noise-traders can therefore create a gap between the price and the fundamental value that will encourage informed traders to pay the cost of acquiring information. The divergence between this price and fundamental value is thereby essential to encourage the process of price discovery. 
%
 \citet[p.529]{BlackNoise} suggested that this combination was essential to the smooth function of financial markets. It is ``the arbitrary element in expectations''. Therefore, noise makes financial markets possible but imperfect: the more noise-traders that there are, the more liquidity and the easier it is to trade; the more noise-traders, the higher the level of inefficiency in the market and the more likely that price will be driven from fundamental value.

@@

However, though the idea that noise is a random variable may be a construction that is useful for modelling and it may be justified in analysis of broader themes of microstructure and levels of market efficiency, this exogenous construct may be an over-simplification that is inappropriate when looking for more understanding of speculation itself.  Institutional and psychological explanations for noise-trading frequently suggest that there are endogenous non-random elements.

The institutional explanations are associated with the allocation of credit for investment and trading purposes:  as asset values decline exchanges, banks and other providers of liquidity require larger margins and deeper haircuts; investors liquidate their fund holdings; the attraction of hedging increases.  See \citet{BrunnermeierLiquidity} and \citet{Gorton} for an overview of the institutional literature.  Psychological explanations revolve around the common panic, the group-thinking and the tendency to read private information from the actions of other investors: the so-called \emph{wisdom of crowds} depends on independent evaluation but as Keynes noted in his discussion of the \emph{beauty contest}, it is generally more important to identify what other investors will believe than actual underlying, fundamental value \citet[p.101]{Keynes1936}

it has also been suggested that speculative techniques can change overtime.  For example, \citet{ShillerFashion} discusses the use of fashionable techniques or fads in investment.  In this way, speculation noise is associated with sentiment and measures of sentiment are then used as ways of measuring speculation. 
%It would be good to have some more modern ideas here.  Maybe they are just the empirical results. 
%this paragraph is not very good.  It needs to be updated.  Look at the file in citulike. 
\subsection{Empirical work}
Empirical assessment of speculation has been limited by the scarcity of data.  It is not very easy to distinguish between speculators and other types of investors.  It is even harder to distinguish those that may be noise-traders from those assumed to be informed.  One approach has been to assume that noise-traders utilise simple investment strategies and to use tests of these strategies as a means of testing speculation.  For example speculation can be identified with momentum trading. It is also possible to assume that certain investors, such as investment banks and hedge funds, are informed while other trade without information. The latter group may include day-traders or private individuals.

@@

The debate about a tax on financial transactions, initiated by \citet{TobinTax} and now called the \emph{the Tobin tax}, has been intended to reduce speculation.  The surge in commodities prices that was seen prior to the GFC was seen to be partially the consequence of increased speculation in commodities and the creation of exchange traded funds (ETF) focusing on commodities markets.  The argument for a tax on financial transactions depends on the idea that speculators will drive the price away from its fundamental value.

@@

\section{Methodology}
The aim of this study is to test whether the noise-trader model provides an appropriate explanation for speculator activity in the  foreign exchange market. This will be use to gain a fuller understanding of speculative behaviour.   The experiment will involve three steps:  assess the price implication of the noise-trader model;  measure speculative activity in the foreign exchange market; finally, assess whether extreme speculative activity in the foreign exchange market is consistent with the noise-trader model.  If the speculative activity that is seen in the foreign exchange market produces results that are like the noise-trader model, it can be said that speculation is noise trading as Keynes asserted, if not it may be more informed as suggested by Friedman.  
% where it is widely perceived that there are a relatively large number of traders engaged in technical analysis.  See \citet gc{FrootChartists} for a survey.  %this may go in the explanation for why fx is used.  Use this elsewhere. Add more?

@@

The following four effects are noted by \citet[pp. 14-15]{Delong1990noise}:  the \emph{Friedman Effect}, as informed speculators move prices back towards fundamental value; \emph{Price Pressure} as the speculative noise-traders drive prices towards their misperceived value; \emph{Hold More} as noise-traders misperceive the value of the risky asset and tend to hold more of it; \emph{Create Space} is the noise-trader risk caused by the difference between their perception and the actual value, it is an added risk that is caused by the variability of noise-trader belief as noise-traders could drive the price even further from its fundamental value. In this model, the more numerous the noise-traders, the more volatile prices.

The Friedman-effect and the Price-pressure are the effects that are studied here.  Their combination suggests that when the weight of noise-traders is high or the times when they have particularly strong opinions will tend to be those when there is the greatest risk of reversals.  The hold-more effect suggests that the weight of speculative involvement in the market will tend to increase as a trend.  Create-space effect would suggest that when the weight of noise-traders is high, there is increased risk.  %this can be assessed.

@@

To understand what the noise-trader model of speculative behaviour means for foreign exchange rates, a \emph{Monte Carlo} simulation is carried out with a variety of misperception-perceptions and noise-trader weights.  Two time series are created.  The first series is based on a random noise-trade misperception, $(\rho^*)$.  This is called the Misperception Model (MPM). The second is based on a random weight of noise-traders $(\mu)$ and is called the Random Weight Model (RWM).   The simulation creates a 1000 period price series from Equation \ref{eq:NP2}. The dividend paid on the risky asset $(r)$ is posted at 10\%; the risk-aversion parameter $(\gamma)$ is assumed to be 0.3 (in line with that suggested in \citet[p. 127]{FrobesBehaviour}); starting from a price of 100, the average noise-trader misconception $(p^*)$ is a random variable in the first simulation with a mean of 4 and a variance $(\sigma_{\rho^*}^2)$ of 2, and has a fixed value of 4 in the second simulation; the weight of noise-traders $(\mu)$ is 50\% in the first case but is modelled as a two-period moving average of a normal random variable centred at 50\% with a standard deviation of 20\% and equally weighted average of its previous value (starting at 50\%) to allow smooth adjustment, in the second simulation.   
% it is possible to have other specifications. For example, the noise-traders oculd become more prominent if their strategy is more successful,

A 1000 period prices simulation is created for each model and the returns for these series are calculated as the log change in price.  Those periods when misperception-perception or the weight of speculators is \emph{extreme} are identified as those that are at or above the ninety-fifth percentile.  These extremes are the \emph{event} for an event study. The R package \emph{eventstudies} created by \citet{eventstudies} is used for the data manipulation.  Descriptive statistics for the whole series of returns and for those when either the extent of their misperception or the weight of noise-traders is extreme are presented in Table \ref{tabref:MC1}.  The process is then repeated for one thousand simulations of each model so that confidence intervals can be established for the estimates of the descriptive statistics.

@@

Table \ref{tabref:DS1} in Appendix \ref{apref:1} shows descriptive statistics for the exchange rate data used in the study.  There are daily data for the period between January 1996 and September 2002.  This is the period for which risk-reversal data are available.  The series are of sightly different lengths as there are some gaps in the risk reversal and exchange rate data as a result of holidays and market closures.  From the raw data continuously compounded returns are calculated and multiplied by 100.  The first thing of note is that the mean return does not appear to be statistically different from zero.  This is what would be expected if, as has been identified in many studies, exchange rates follow a random walk. 
\begin{sidewaystable}
\begin{threeparttable}
\caption{Descriptive Statistics for Risk Reversals: 2 January 1996 to 6 September 2002}
\begin{tabular}{l | r r  r r r r r r r r}
  \hline
 Statistic &  AUD & EURGBP & EURSEK & GBP & CHF & EURCHF & EURJPY & EUR & CAD & JPY \\
 \hline
Number & 1702 & 1697 & 1719 & 1691 & 1374 & 1339 & 1697 & 1692 & 1701 & 1697 \\ 
  Mean & -0.37 & 0.15 & 0.71 & -0.08 & -0.28 & -0.38 & -0.48 & 0.19 & 0.12 & -0.74 \\ 
  Median & -0.30 & 0.10 & 0.80 & -0.10 & -0.15 & -0.30 & -0.10 & 0.10 & 0.15 & -0.55 \\ 
  Standard Deviation & 0.38 & 0.37 & 0.51 & 0.33 & 0.49 & 0.36 & 1.02 & 0.49 & 0.25 & 1.02 \\ 
 Max & 0.45 & 1.72 & 1.90 & 1.00 & 1.05 & 0.50 & 1.70 & 2.05 & 1.45 & 2.05 \\ 
 Min & -1.70 & -0.80 & -0.30 & -1.35 & -2.05 & -1.80 & -4.03 & -1.15 & -0.65 & -4.00 \\ 
 Skewness & -0.50* & 1.27* & -0.23* & 0.33* & -0.79* & -0.58* & -0.93* & 0.68* & 0.23* & -0.51* \\ 
  Kurtosis & 0.02 & 2.29* & -1.67 & 0.72 & 0.69 & 0.19 & 0.15 & 0.96 & 1.83 & -0.03 \\ 
      \hline
 \label{tabref:DS2}
\end{tabular}
\begin{tablenotes}
\small
\item Daily risk reversal skew.  Risk reversal is the premium of implied volatility for calls relative to the equivalent put (equal distance to the strike price, 25 delta in this case).  A figure above zero means that calls are more expensive (in volatility and price terms) than puts.  A figure below zero means that puts are more expensive. The asterisk (*) means that the estimate is significant at the 5\% level using t-statistics for the mean and the standard error of skew and standard error of excess-kurtosis respectively.   These are calculated as $SES = \sqrt{\frac{6n(n-1)}{(n-2)(N+1)(N+3)}}$ and $SEK = \sqrt{\frac{n^2-1}{(n-3)(N+5)}}$, with the assumption that this estimate is distributed normally, the skew or kurtosis is considered significant if it is more than two standard errors from zero.
\end{tablenotes}
\end{threeparttable}
\end{sidewaystable}

@@

\begin{sidewaystable}
\begin{threeparttable}
\caption{Descriptive Statistics for S1 (Speculative Sentiment) and S2 (Speculative Weight): 30 September 1998 to 31 December 2008}
\begin{tabular}{l | r r r r r r r  r r r }
  \hline
 Statistic &  \multicolumn{2}{c}{EUR} & \multicolumn{2}{c}{JPY} & \multicolumn{2}{c}{GBP} & \multicolumn{2}{c}{CHF} & \multicolumn{2}{c}{CAD}  \\
 & S1 & S2 & S1 & S2 & S1 & S2 & S1 & S2 & S1 & S2 \\  
\hline
Number & 522 &  522 & 849 & 849 & 849 & 849 & 847 & 847 & 849 & 849\\
Mean &  31.31 & 43.43 & -23.90 & 44.21 & 2.76 & 41.68 & -21.14 & 46.13 & 6.68 & 41.23\\
Median & 42.75 & 39.56 & -33.30 & 43.56 & 11.15 & 38.88 & -37.32 & 42.88 & 7.44 & 39.81\\
Standard deviation  & 41.94 & 16.13 & 51.49 & 12.17 & 60.59 & 16.46 & 60.76 & 15.73 & 55.26 & 15.33\\
Max & 95.67 & 86.06 & 87.43 & 84.50 & 96.15 & 89.45 & 100.00 & 92.49 & 100.00 & 86.63\\
Min &  -100.00 & 11.15 & -99.83 & 12.88 & -100.00 & 9.10 & -100.00 & 14.61 & -90.08 & 11.89\\
Skew  & -0.88* & 0.55* & 0.33* & 0.57* & -0.17 & 0.76* & 0.45* & 0.65* & 0.02 & 0.50*\\
Kurtosis & 0.18 & -0.57 & -1.14 & 0.64 & -1.45 & 0.15 & -1.25 & -0.15 & -1.35 & -0.28\\
\hline
 \label{tabref:DS4}
\end{tabular}
\begin{tablenotes}
\small
\item S1 is the first sentiment indicator, calculated as ratio of net non-commercial long less non-commercial short to the sum of non-commercial long and short positions multiplied by one hundred.  S2 is the second sentiment indicators, calculated as the ratio of net non-commercial long to net non-commercial short plus spreading to open interest.   An asterisk (*) means that the estimate is significantly different from zero at the 5\% level using t-statistics for the mean and the standard error of skew and standard error of excess-kurtosis respectively.  These are calculated as $SES = \sqrt{\frac{6n(n-1)}{(n-2)(N+1)(N+3)}}$ and $SEK = \sqrt{\frac{n^2-1}{(n-3)(N+5)}}$, with the assumption that this estimate is distributed normally, the skew or kurtosis is considered significant if it is more than two standard errors from zero.
\end{tablenotes}
\end{threeparttable}
\end{sidewaystable}

\subsection{Analysis of events and abnormal returns}
The aim of the study is to assess the effect of extreme speculative positions on future returns to ascertain the informational content of these extremes.  Therefore, the events that will be used to assess abnormal returns are the extreme readings on sentiment and the extreme presence of speculators.   Any reading above the ninety-ninth percentile or below the first percentile is deemed to be extreme.  For the risk reversal skew, as a secondary test, a reading above the ninety-fifth and below the fifth percentile is also established as an event.   The \citet{eventstudies} R package was used to manage the data.  The cumulative abnormal returns for the whole of the event window (CARW) and the cumulative abnormal returns for after the event (CARA) were recorded.  These cumulations are just the summation of the continuously compounded returns over the appropriate period.

@@

What is more interesting is whether knowledge about the extreme reading at time $t = 0$ provides any information about future returns.  The Friedman-effect would suggest a reversal after an extreme.   This information is provided in columns 5, 8, 11 and 14 of Table \ref{tabref:RR1}. A reversal is indicated by a significant negative reading for highs or positive reading for lows.   Looking, for example, at the four-day window in column 8, it is clear that the extreme risk reversal skew does not provide any information about future returns.  Of the eighteen cases using the 95\% th percentile to identify an extreme, ten show a continuation, rather than a reversal, that is not significantly different from zero; one shows a significant continuation; five show reversals that are not significant; only two show significant reversal (EURCHF in each case).  For the more extreme event when the 99\% the percentile is used to measure the extreme, there is an even more mixed picture.  Just over half the cases with the four-day window show a reversal (8 of 18), but only 2 of these are statistically significant.

\begin{sidewaystable}

@@

@@

\end{sidewaystable}

@@

Once again, there is a very strong association between returns and the movement of speculators.  For the 4-day window and the S1 and S2 measures, every one of the currencies studied, high and low, had a mean cumulative return for the whole of the event window that was in the same direction as sentiment.  When re-sampled a thousand times, more than ninety five percent of the means calculated were greater than zero in all cases but one (GBP).  Once again there is evidence here of momentum and of speculator positions following the movement of prices. This consistent with both the gradual absorption of information into the price and with the feedback of price into additional speculation.

However, there is little evidence that these speculators are driving prices to extreme deviations from fundamentals with their extreme behaviour.  Sixty percent of the S1 cases still show a significant abnormal return in the direction of the extreme 4 weeks after the extreme has been reached; only thirty percent of the cases show a reversal and in only one of those do ninety five percent of the re-sampled means remain above zero.  For S2 there is even less ambiguity, with ninety percent of the cases showing a continuation and forty percent of them being significant.  Again the evidence is consistent with the gradual adjustment of prices to information.
